The impact of gas bulk rotation on the Lyman-$\alpha$ line

### 2013-09-05 paper/Analisis.tex
one more couple of references
a couple of paragraphs in the introduction. added new references

A more detailed quantification of the \ly line has been reached using
Monte Carlo simulations \citep{Auer68,Avery68,Adams72}. In the last
two decades these studies have become more popular due to the
availability of computing power. Early into the 21st century the first
studies focused on on homogeneous and static media
\citep{Ahn00,Ahn01,Zheng02}. Later on the effects of clumpy media
\citep{Hansen06} and expanding/contracting shell/spherical geometries started to
be studied \citep{Verhamme06,Dijkstra06}. Similar codes have applied
these results to semi-analytical models of galaxy formation \citep{Orsi12} and
results of large hydrodynamical simulations
\citep{CLARA,Forero-Romero12}. Recently Monte Carlo codes have also
been applied to the results of high resolution hydrodynamical
simulations \citep{Laursen09,Yajima12}.

. In this
context a lot of attention has been paid to effects that might induce
asymmetric line profiles and to the estimation of the escape fraction in the presence of dust.

The bulk gas velocity
plays a central role in the theoretical studies of the \ly line.

[There is an extensive literature studying the influence of
outflow/inflow configurations in the shape of the outgoing Lyman-alpha
line.]

[Rotation is present in high redshift galaxies.]

In this paper we study for the first time the impact of rotation on
the morphology of the Lyman $\alpha$ line. To isolate the effects of
rotation we focus on a simple system: the gas distribution is
spherical, with homogeneous density and the gas rotates as a solid
body. We base our work on two independent Monte Carlo based radiative
transfer codes CLARA \citep{CLARA} and XX \citep{DijkstraKramer}.

>>>>

A more detailed quantification of the \ly line has been reached using
Monte Carlo simulations \citep{Auer68,Avery68,Adams72}. In the last
two decades these studies have become more popular due to the
availability of computing power. Early into the 21st century the first
studies focused on on homogeneous and static media
\citep{Ahn00,Ahn01,Zheng02}. Later on the effects of clumpy media
\citep{Hansen06} and expanding/contracting shell/spherical geometries started to
be studied \citep{Verhamme06,Dijkstra06}. Similar codes have applied
these results to semi-analytical models of galaxy formation \citep{Orsi12} and
results of large hydrodynamical simulations
\citep{CLARA,Forero12,Behrens13}. Recently Monte Carlo codes have also
been applied to the results of high resolution hydrodynamical
simulations \citep{Laursen09,Barnes11,Verhamme12,Yajima12}. Recent developments have been
focused on the study of clumpy outflows \citep{DijkstraKramer}and
anisotropic velocity configurations \citep{Zheng2013}.

The recent studies of galaxies in hydrodynamical simulations
\citep{Laursen09,Barnes11,Verhamme12,Yajima12} have all shown
systematic variations in the \ly line with the viewing angle. These
variations are a complex superpositions of anisotropic density
configurations (i.e. edge-on vs. face-on view of a galaxy), the
inflows observaed by gas cooling and the outflows included in the
supernova feedback process of the simulation. These bulk flows
physically correspond to the circumgalactic and intergalactic medium
(CGM and IGM). These effects have been systematically studied
different anisotropic models of hydrogen clouds  varying the density
and wind characteristics \citep{Zheng2013}.

However, in all these systematics efforts the effect of rotation, which is an ubiquitous
feature in galaxies, has not been systematically studied. The
processing of the \ly photons in a rotating interstellar medium (ISM)
must have an impact in the line morphology. It is necessary to investigate to
what extent bulk gas rotation has an measurable impact in the \ly
alpha line.

Performing that study is the main goal of this paper. We study for the
first time the impact of rotation on the morphology of the Lyman
$\alpha$ line. We focus on a simplified system, a spherical gas claud
with homogeneous density and solid body rotation. We base our work on two independent Monte
Carlo based radiative transfer codes CLARA \citep{CLARA} and XX
\citep{DijkstraKramer} .
